Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy

### 2023-11-20 manuscript/manuscript.qmd
Update word count

---
title: "Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy"
short-title: "Why Donors Donate"
date: "2023-11-18"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
  orcid: 0000-0003-3506-7590
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
  orcid: 0000-0001-7846-3660
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  Non-governmental organizations (NGOs) are a core component of global governance. Over the last two decades, however, many host governments have cracked down on NGOs using a series of laws that adversely affect the availability and use of traditional NGO funding. In response to this dramatic shift in the funding landscape, international NGOs have increasingly turned to individual donors to offset the loss funding. Prior research on NGO fundraising has examined how financial transparency and organizational accountability influence individual donor behavior. How do these elements of private donor decision-making interact with structural-level factors such as worsening host country civic environments when deciding to donate? Using a conjoint experiment with likely donors in the US, we find that transparency and accountability can protect against the dampening effect of host government crackdown and criticism of NGOs. Our results have important implications for how NGOs can adapt to worsening civic environments. 
keywords:
  - philanthropy
  - civil society
  - conjoint experiments
  - donor heuristics
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/silent-skywalk>"
bibliography: references.bib
csl: csl/british-journal-of-political-science.csl
link-citations: true
word-count: 9350

>>>>

---
title: "Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy"
short-title: "Why Donors Donate"
date: "2023-11-18"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
  orcid: 0000-0003-3506-7590
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
  orcid: 0000-0001-7846-3660
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  Non-governmental organizations (NGOs) are a core component of global governance. Over the last two decades, however, many host governments have cracked down on NGOs using a series of laws that adversely affect the availability and use of traditional NGO funding. In response to this dramatic shift in the funding landscape, international NGOs have increasingly turned to individual donors to offset the loss funding. Prior research on NGO fundraising has examined how financial transparency and organizational accountability influence individual donor behavior. How do these elements of private donor decision-making interact with structural-level factors such as worsening host country civic environments when deciding to donate? Using a conjoint experiment with likely donors in the US, we find that transparency and accountability can protect against the dampening effect of host government crackdown and criticism of NGOs. Our results have important implications for how NGOs can adapt to worsening civic environments. 
keywords:
  - philanthropy
  - civil society
  - conjoint experiments
  - donor heuristics
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/silent-skywalk>"
bibliography: references.bib
csl: csl/british-journal-of-political-science.csl
link-citations: true
word-count: 7912
